Large scale risks from pesticides in small streams

### 2016-10-19 article.tex
more fixes, typos etc

\usepackage{todonotes}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% 150-200 words
Small streams are important refugia for biodiversity.
In agricultural areas they may be at high risk from pesticide pollution. 
However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales. 
We used data from German governmental water quality monitoring to quantify the drivers of pesticide risk and to assess pesticide risk in small streams on a large scale. 
The data set comprised of 1,766,104 measurements related to 24,743 samples in 2,301 sampling sites and of 478 pesticides.  
We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics of pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification. 
Agricultural land use lead to a 3.4-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent. 
Precipitation increased measured pesticide risk by 5\% and was thw highest during summer months.
Risk thresholds were exceeded in 26 percent of small streams, with the highest risk emerging from neonicotinoid insecticides. 
We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events.

@@

Small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
Nevertheless, a clear definition of SS in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
For example, the WFD defines SS with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2} and \citet{lorenz_specifics_2016} defines SS with catchment size \textless 10km\textsuperscript{2}.

@@

In this study we compiled and analyzed large-scale chemical monitoring data from small streams in Germany. 
First, we analysed the shape of the relationship between pesticide risk, agricultural land use and catchment size and examined whether related thresholds for pesticide risks can be derived. 
Second, we investigated the influence of precipitation and seasonal dynamics on pesticide detections, given that precipitation proved an important driver of pesticide exposure in several small-scale studies \citep{wittmer_significance_2010}\citep{schulz_field_2004}, but it is unknown whether a precipitation signal prevails on large scales. 
Finally, we quantified the current risks from pesticides in small streams in Germany.

%% -------------------------------------------------------------------------
\section{Methods}
\subsection{Data compilation}
We requested pesticide monitoring data from sampling sites that can be classified as small streams (catchment sizes $\mathrm{< 100km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names). 
The request encompassed only the years 2005 to 2015.
We homogenized and unified all data provided by the federal states into a database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.

@@

For each derived catchment (either from DEM or drainage basins) we calculated the percentage agricultural land-use within the catchment based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}. 
Thus, agricultural land use information was available for 98\% of all sites (24\% from authorities, 52\% from DEM and 22\% from drainage basins). 
For 97\% of the sites both, the proportion of agricultural land use and catchment size were available. 
68 sites (3\%) lacking these two informations were omitted from the operations outlined below.
%see do_overview.R

@@

We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
We modeled the number of RAC exceedances at a site as:

@@

where $No(RQ > 1)_i$ is the observed number of RAC exceedances at site $i$. 
We modeled $No(RQ > 1)_i$ as resulting from a negative binomial distribution ($NB$) with mean $\mu_i$ and a quadratic mean-variance-relationship ($Var(No(RQ > 1)_i) = \mu_i + \frac{\mu_i^2}{\kappa}$).
The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of the number of RAC exceedances. 
$f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006} and $\beta_0$ is the intercept.
The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modeling the rate of exceedances. 
We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
GAMs were fitted using the mgcv package \citep{wood_fast_2011}.

@@

To account for temporal auto-correlation and differences between federal states we used $site$ nested within $state$ as random intercepts.
We implemented this model using the gamlss package \cite{stasinopoulos_generalized_2007}.

We fitted this model separately to each compound with a RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 23 compounds, see Supplemental Table~S3 for a list of compounds). 
To summarise the coefficients across the 24 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 24 compounds.
The results of individual compounds are provided in the Supplemental Table~S4 and Figure~S7.
The meta-analysis was performed using the metafor package \citep{Viechtbauer_2010}.

@@

The compiled dataset used in analysis comprised 1,766,104 pesticide measurements of 24,743 samples at 2,301 sampling sites in small streams.  %see do_overview.R for numbers.
These samples were all taken via grab sampling.  
We found large differences in the number of sampling sites between federal states and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1). 
The number of small stream sampling sites per state ranged from 1 (Lower Saxonia, NI) to 1139 (North Rhine-Westphalia, NW) and no data from Brandenburg.

\begin{figure}[ht]
  \includegraphics[width=0.6\textwidth]{figure1.pdf}
  \caption{Spatial distribution of the 2,301 small stream sampling sites. Colour codes different federal states, see Supplemental Table~S1 for abbreviations.}
  \label{fig:fig1}
\end{figure}

In total 478 different compounds used as pesticides and their metabolites were measured at least once (Supplemental Table~S2). 
Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109). % see overview.R
Most samples were taken in the months April till October, with fewer samples during winter (see Supplemental Figure~S3).
4\% (71,113) of all measurements were detects above LOQ.
We found substantial differences in the spectra of analyzed pesticides between federal states (Figure~\ref{fig:fig2}).
The number of different pesticides per state ranged from 57 (SL) to 236 (RP) (Supplemental Table~S1).
Hierarchical clustering revealed that RP and NI analysed a distinct compound spectra compared to the cluster of other states.
Although there was high variation within the remaining cluster, this could not be further split (Figure~\ref{fig:fig2}, also Supplemental Figures~S4 and S5).

@@

$prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
An increase of precipitation in $Q2$ before sampling ($prec_{-1}$) from 1mm to 10mm precipitation lead to a 5\% higher mean RQ and the probability to exceed LOQ increases from 4\% to 6\% on average (Figure~\ref{fig:fig5}, top). % do_precip.R
Effects on individual compounds are provided in the Supplemental Table~S4.
Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$). 
This difference was less pronounced for the mean value of RQ (Figure~\ref{fig:fig5}, top).

@@

\begin{figure}[ht]
  \includegraphics[width=0.8\textwidth]{figure5.pdf}
  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter (bottom row) from a meta-anaylsis of the 24 modeled compounds. Left column: coefficients for mean RQ ($\mu$), right column: coefficients for probabilty to exceeed LOQ ($\nu$). 
  Coefficients are shown on the link scale (see Eq.~\ref{eqn:eqn4}).
  Single compound coefficients are provided in the Supplemental Table~S4 and Figure~S7).
  }
  \label{fig:fig5}
\end{figure}

\subsection{Pesticide risk in small streams}
We found RQ > 1 in 26\% of sampling sites and RQ > 0.1 in 54\% of sites. 
In 23\% of sites chemicals with RAC were never detected (see also Supplemental Figure~S8).
Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ (Figure~\ref{fig:fig6}). %do_pollution.R
For Thiacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have an RQ~\textgreater~1. 
The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (26.7, 14.1 and 21.1 \% of measurements~\textgreater~LOQ).
RAC exceedances were found in 15\% of samples with detects (and 7.7\% of all samples with RAC).

The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Clothianidin (max(RQ) = 157), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 84\% of all samples were it was analysed (n = 3038, see also Supplemental Figure~S9).
Glyphosate was the compound with the highest detection rates (41\%, n = 3557 samples), followed by Boscalid (23\%, n = 9886) and Isoproturon (22\%, n = 19112). 
However, only the latter showed RQ exceedances (Figure~\ref{fig:fig6}).
In 45.9\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8).

@@

%% -------------------------------------------------------------------------
\section{Discussion}
\subsection{Overview on the compiled dataset}
The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive one available for Germany.
Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States \citep{stone2014pesticides}.
While the compilations from Europe are of similar quantity and quality to the  data compiled and analysed here for Germany, the compilation used in \citet{stone2014pesticides} is much smaller.
Nevertheless, there might more data for the United States available (e.g. from Water Quality Portal (WQP), \url{www.waterqualitydata.us}).

% current problems in monitoring and possible solutions
A nation-wide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states. 
Although, the outlined differences between states, all ecoregions occurring Germany \citep{illies1978limnofauna,abell2008freshwater} are covered by the presented dataset and therefore, may represent a sample of all small streams in Central Europe. 
For Thiacloprid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
For these compounds a lowering of LOQ is essential for reliable assessment.
Moreover, a nation-wide assessment would benefit from a harmonized spectrum of analysed compounds between federal states.

Given their high abundance in the landscape \citep{nadeau_hydrological_2007} small streams are underrepresented in the current monitoring (Figure~\ref{fig:fig3}). 
Especially, sampling sites in streams below 10~km\textsuperscript{2} are disproportionally low, which may be attributed to the missing categorisation in the WFD. 
Clearly, there is currently a lack of knowledge for these small streams.
We analysed only data from small streams, however, for lentic small water bodies this lack might be even greater \citep{lorenz_specifics_2016}.

\subsection{Influence of agricultural land use and catchment size}
% agriculture
We found a strong influence of agriculture on the pollution of streams.
If there is more the 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase in fully agricultural catchments (above 75 \% agriculture).
To our knowledge, this is the first study investigating such thresholds of pesticide risk.
Previous studies examined thresholds for percent agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
\citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use in lowland streams in Europe.
Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\% agricultural land use in wadeable streams in the United States.
Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological changes.

@@

We found highest the probability of exceeding LOQ during summer (10\% for Q2) and lowest in the first quarter (4\%, Figure~\ref{fig:fig5}, bottom right).
This yearly pattern coincides with their main application season for pesticides in Central Europe.
Nevertheless, there are compound specific differences in the yearly pattern, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
For example, the herbicide Diflufenican showed highest RQ and probability of exceeding LOQ during the winter quarters Q1 and Q4 (Supplemental Table~S4), which coincides with the application period it is registered for in Germany \citep{bvl_online_2016}.
Our study suggests that pesticide risks display compound specific spatio-temporal dynamics.
Currently, little is known about these and further research on those might provide useful information for future ecological risk assessment. 
For example, the sensitivity of organisms is often life stage dependent \citep{hutchinson1998analysis} and knowledge on temporal dynamics could inform on concurrent exposure to multiple pesticides, as well as assist to parameterise toxicokinetic and toxicodynamic models \citep{ashauer2016modelling}.

\subsection{Pesticides in small streams}
Our results suggest that small streams are frequently exposed to ecologically relevant pesticide concentrations.
In one quarter of small streams RACs were exceeded at least once.
\citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
By contrast, we found that that neonicotinoid insecticides have highest exceedances of RACs, followed by the organophosphate chlorpyrifos. 
This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (for example 3,540 samples of Thiacloprid, Figure~\ref{fig:fig6}). 
Overall, our results suggest that neonicotinoids may currently pose a high risk to freshwater ecosystems. 
Our results add further evidence to the growing literature on risks arising from neonicotinoids for aquatic \citep{morrissey2015neonicotinoid} and terrestrial ecosystems \citep{pisa2015effects}.

@@

By contrast, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
However, RAC exceedances were generally lower and less abundant compared to our results from Germany. 
This might reflect differences in pesticide use between countries and ecoregions. 
From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
Indeed, \citet{stehle_agricultural_2015} found that biological diversity of stream invertebrates was significantly reduced by 30\% at RQ = 1.12 and by 10\% at 1/10 of RAC.
We found RQ values greater than 1.12 in one quarter of small streams and RQ at 1/10 of RAC in more the half of streams. 
Consequently, we conclude that on a large-scale agricultural pesticides are a major threat to small streams, the biodiversity they host and the services they provide. 
This threat may exacerbate because pesticides often occur in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}
The authors thank the federal state authorities for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding a related project (FKZ 3714 67 4040 / 1). 
\end{acknowledgement}

%%% Word count
% abstract    :  200
% 4 big(600)  : 2400
% 2 small(300):  600
% acknowledg  :   30
%              =====
% subtotal      3227
% text body   : 3700
% ======================================
% total       : 6930

>>>>

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% 150-200 words
Small streams are important refugia for biodiversity.
In agricultural areas they may be at high risk from pesticide pollution. 
However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales. 
We used data from German governmental water quality monitoring to quantify the drivers of pesticide risk and to assess pesticide risk in small streams on a large scale. 
The data set comprised of 1,766,104 measurements related to 24,743 samples in 2,301 sampling sites of 478 pesticides.  
We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics of pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification. 
Agricultural land use lead to a 3.4-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent. 
Precipitation increased measured pesticide risk by 5\% and was the highest during summer months.
Risk thresholds were exceeded in 26 percent of small streams, with the highest risk emerging from neonicotinoid insecticides. 
We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity and that governmental pesticide sampling should be adapted to precipitation events.

@@

Small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
Nevertheless, a clear definition of small streams in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
For example, the WFD defines small streams with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2} and \citet{lorenz_specifics_2016} defines small streams with catchment size \textless 10km\textsuperscript{2}.

@@

In this study we compiled and analyzed large-scale chemical monitoring data from small streams in Germany. 
First, we analysed the shape of the relationship between pesticide risk, agricultural land use, and catchment size and examined whether related thresholds for pesticide risks can be derived. 
Second, we investigated the influence of precipitation and seasonal dynamics on pesticide detections, given that precipitation proved an important driver of pesticide exposure in several small-scale studies \citep{wittmer_significance_2010}\citep{schulz_field_2004}, but it is unknown whether a precipitation signal prevails on large scales. 
Finally, we quantified the current risks from pesticides in small streams in Germany.

%% -------------------------------------------------------------------------
\section{Methods}
\subsection{Data compilation}
We requested pesticide monitoring data from sampling sites that can be classified as small streams (catchment sizes $\mathrm{< 100km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names). 
The request encompassed only the years 2005 to 2015.
We homogenized and unified all data provided by the federal states into a database and implemented a robust data-cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.

@@

For each derived catchment (either from DEM or drainage basins) we calculated the percentage agricultural land-use within the catchment based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}. 
Thus, agricultural land use information was available for 98\% of all sites (24\% from authorities, 52\% from DEM and 22\% from drainage basins). 
For 97\% of the sites both, the proportion of agricultural land use and catchment size were available. 
68 sites (3\%) lacking these two pieces of information were omitted from the operations outlined below.
%see do_overview.R

@@

We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
We modeled the number of RAC exceedances (RQ \textgreater 1) at a site as:

@@

where $No(RQ > 1)_i$ is the observed number of RAC exceedances at site $i$. 
We modeled $No(RQ > 1)_i$ as resulting from a negative binomial distribution ($NB$) with mean $\mu_i$ and a quadratic mean-variance-relationship ($Var(No(RQ > 1)_i) = \mu_i + \frac{\mu_i^2}{\kappa}$).
The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of the number of RAC exceedances. 
$\beta_0$ is the intercept and $f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}. 
The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modeling the rate of exceedances. 
We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
GAMs were fitted using the mgcv package \citep{wood_fast_2011}.

@@

To account for temporal autocorrelation and differences between federal states we used $site$ nested within $state$ as random intercepts.
We implemented this model using the gamlss package \cite{stasinopoulos_generalized_2007}.

We fitted this model separately to each compound with an RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 23 compounds, see Supplemental Table~S3 for a list of compounds). 
To summarise the coefficients across the 24 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 24 compounds.
The results of individual compounds are provided in the Supplemental Table~S4 and Figure~S7.
The meta-analysis was performed using the metafor package \citep{Viechtbauer_2010}.

@@

The compiled dataset used in analysis comprised 1,766,104 pesticide measurements of 24,743 samples at 2,301 sampling sites in small streams.  %see do_overview.R for numbers.
These samples were all taken via grab sampling.  
We found large differences between federal states in the number of sampling sites and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1). 
The number of small stream sampling sites per state ranged from 1 (Lower Saxonia, NI) to 1139 (North Rhine-Westphalia, NW).
No data was available from Brandenburg.

\begin{figure}[ht]
  \includegraphics[width=0.6\textwidth]{figure1.pdf}
  \caption{Spatial distribution of the 2,301 small stream sampling sites. Colour codes different federal states (see Supplemental Table~S1 for abbreviations).}
  \label{fig:fig1}
\end{figure}

In total 478 different compounds used as pesticides and their metabolites were measured at least once (Supplemental Table~S2). 
Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109). % see overview.R
Most samples were taken in the months April till October, with fewer samples during winter (see Supplemental Figure~S3).
4\% (=71,113) of all measurements were concentrations above LOQ.
We found substantial differences in the spectra of analyzed pesticides between federal states (Figure~\ref{fig:fig2}).
The number of different pesticides per state ranged from 57 (SL) to 236 (RP) (Supplemental Table~S1).
Hierarchical clustering revealed that RP and NI analysed a distinct compound spectra compared to the cluster of other states.
Although there was high variation within the remaining cluster, this could not be further split (Figure~\ref{fig:fig2}, also Supplemental Figures~S4 and S5).

@@

$prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
In $Q2$ an increase of precipitation before sampling ($prec_{-1}$) from 1mm to 10mm lead on average to a 5\% higher mean RQ and the probability to exceed LOQ increases from 4\% to 6\% (Figure~\ref{fig:fig5}, top). % do_precip.R
Effects differed between individual compounds and are provided in the Supplemental Table~S4.
Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$) on the probability of exceeding LOQ. 
This difference was less pronounced for the mean value of RQ (Figure~\ref{fig:fig5}, top).

@@

\begin{figure}[ht]
  \includegraphics[width=0.8\textwidth]{figure5.pdf}
  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter (bottom row) from a meta-anaylsis of the 24 modeled compounds. Left column: coefficients for mean RQ ($\mu$), right column: coefficients for probabilty to exceeed LOQ ($\nu$). 
  Coefficients are shown on the link scale (see Eq.~\ref{eqn:eqn4}).
  Single compound coefficients are provided in Supplemental Table~S4 and Figure~S7).
  }
  \label{fig:fig5}
\end{figure}

\subsection{Pesticide risk in small streams}
We found RAC exceedances in 26\% of sampling sites and RQ > 0.1 in 54\% of sites. 
In 23\% of sites no chemicals with RAC were detected (see also Supplemental Figure~S8).
Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ (Figure~\ref{fig:fig6}). %do_pollution.R
For Thiacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have an RQ~\textgreater~1. 
The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (26.7, 14.1 and 21.1 \% of measurements~\textgreater~LOQ), see also Supplemental Table~S5).
RAC exceedances were found in 15\% of samples with concentrations \textgreater LOQ (and 7.7\% of all samples).

The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Clothianidin (max(RQ) = 157), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 84\% of all samples where it was analysed (n = 3038, see also Supplemental Figure~S9).
Glyphosate was the compound with the highest detection rates (41\%, n = 3557 samples), followed by Boscalid (23\%, n = 9886) and Isoproturon (22\%, n = 19112). 
However, only the latter showed RAC exceedances (Figure~\ref{fig:fig6}).
In 45.9\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8).

@@

%% -------------------------------------------------------------------------
\section{Discussion}
\subsection{Overview on the compiled dataset}
The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive one available for Germany.
Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States \citep{stone2014pesticides}.
While the compilations from Europe are of similar quantity and quality to the  data compiled and analysed here, the compilation used in \citet{stone2014pesticides} is much smaller.
Nevertheless, there might more data for the United States available (e.g. from Water Quality Portal (WQP), \url{www.waterqualitydata.us}).

% current problems in monitoring and possible solutions
A nation-wide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states. 
Although the outlined differences between states, all ecoregions occurring in Germany \citep{illies1978limnofauna,abell2008freshwater} were covered by the presented dataset and therefore, may represent a sample of all small streams in Central Europe. 
For Thiacloprid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
For these compounds a lowering of LOQ is essential for reliable assessment.
Moreover, a nation-wide assessment would benefit from a harmonized spectrum of analysed compounds between federal states.

Given their high abundance in the landscape \citep{nadeau_hydrological_2007} small streams are underrepresented in the current monitoring. 
Given their abundance, small streams below 10~km\textsuperscript{2} are disproportionally less sampled (Figure~\ref{fig:fig3}), which may be attributed to the missing categorisation in the WFD. 
Clearly, there is currently a lack of knowledge for these small streams.
We analysed only data from small streams, however, for lentic small water bodies this lack might be even greater \citep{lorenz_specifics_2016}.

\subsection{Influence of agricultural land use and catchment size}
% agriculture
We found a strong influence of agriculture on the pollution of streams.
If there is more the 25\% agriculture within a catchment pesticides, it is likely that an RAC will be exceeded, with a further increase in fully agricultural catchments (above 75 \% agriculture).
To our knowledge, this is the first study investigating such thresholds of pesticide risk.
Previous studies examined thresholds for percent agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
\citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use in lowland streams in Europe.
Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\% agricultural land use in wadeable streams in the United States.
Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological changes.

@@

We found highest the probability of exceeding LOQ during summer (10\% for Q2) and lowest in the first quarter of the year (4\%, Figure~\ref{fig:fig5}, bottom right).
This yearly pattern coincides with their main application season for pesticides in Central Europe.
Nevertheless, there are compound specific differences in the yearly pattern, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
For example, the herbicide Diflufenican showed the highest RQ and the highest probability of exceeding LOQ during the winter quarters Q1 and Q4 (Supplemental Table~S4), which coincides with the application period it is registered for in Germany \citep{bvl_online_2016}.
Our study suggests that pesticide risks display compound specific spatio-temporal dynamics.
Currently, little is known about these and further research on those might provide useful information for future ecological risk assessment. 
For example, the sensitivity of organisms is often life stage dependent \citep{hutchinson1998analysis} and knowledge on temporal dynamics could inform on concurrent exposure to multiple pesticides, as well as assist to parameterise toxicokinetic and toxicodynamic models \citep{ashauer2016modelling}.

\subsection{Pesticides in small streams}
Our results suggest that small streams are frequently exposed to ecologically relevant pesticide concentrations.
In one quarter of small streams RACs were exceeded at least once.
\citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
By contrast, we found that that neonicotinoid insecticides have highest exceedances of RACs, followed by the organophosphate chlorpyrifos. 
This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (for example 3,540 samples of Thiacloprid, Figure~\ref{fig:fig6}). 
Overall, our results suggest that neonicotinoids may currently pose a high risk to freshwater ecosystems. 
Our results add further evidence to the growing literature on risks arising from neonicotinoids for aquatic \citep{morrissey2015neonicotinoid} and terrestrial \citep{pisa2015effects} ecosystems.

@@

By contrast, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
Moreover, RAC exceedances in Switzerland were generally lower and less abundant compared to our results from Germany. 
This might reflect differences in pesticide use between countries and ecoregions. 
From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
Indeed, \citet{stehle_agricultural_2015} found that biological diversity of stream invertebrates was significantly reduced by 30\% at RQ = 1.12 and by 10\% at 1/10 of RAC.
We found RQ values greater than 1.12 in one-quarter of small streams and RQ at 1/10 of RAC in more then half of small streams. 
Consequently, we conclude that agricultural pesticides are on a large scale a major threat to small streams, the biodiversity they host and the services they provide. 
This threat may exacerbate because pesticides often occur in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}
The authors thank the federal state authorities for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding a related project (FKZ 3714 67 4040 / 1). 
We thank Wolfram König and Alexandra Müller (German Federal Environmental Protection Agency (UBA)), Martin Keller and Beate Bänsch-Baltruschat (German Federal Institute of Hydrology (BfG)), Matthias Liess and Kaarina Foit (Center for Environmental Research (UFZ)) for their valuable contributions to this project. 
\end{acknowledgement}

%%% Word count
% abstract    :  200
% 4 big(600)  : 2400
% 2 small(300):  600
% acknowledg  :   70
%              =====
% subtotal      3270
% text body   : 3700
% ======================================
% total       : 6970
